---
title: "Figure1,2 Boxplots"
output:
  pdf_document: default
  html_notebook: default
author: "Maha Ramamurthy" 
Date: "June 2024"
Place: "Stanford university"
---

```{r}

library(rstatix)

library(plotrix)
library(dplyr)
library(tidyr)
library(tidyverse)
library(psych)
library(cocor)
library(stargazer)
library(data.table)
library(gtsummary)
library(kableExtra)
library(here)
library(Boruta)
library(gtsummary)
library(ggpubr)
library(data.table)
library(ggExtra)
library (psych)
library(plotROC)
library(rstatix)
library(plotrix)
library(dplyr)
library(tidyr)
library(tidyverse)
library(psych)
library(stargazer)
library(gtsummary)
library(ggpubr)
library(ggExtra)
library(cutpointr)
library(lme4)
library(dplyr)
library(ggplot2)
library(effsize)
```

```{r}
#Read all the dfs with demographic and reading outcome measures 
df_demo_LET_Available <- read.csv("~/LET.csv")
df_demo_PSE_Available <- read.csv("~/PSE.csv")
df_demo_GMC_Available <- read.csv("~/GMC.csv")
```

# How correlated are primary Language and FRPM?
```{r}

table(df_demo_LET_Available$ELStatus, df_demo_LET_Available$proxy_SES)
# These are dichotomous variables and therefore need to report the phi coefficient not really r! 
cor(ifelse(df_demo_LET_Available$ELStatus=="EO", 0, 1),ifelse(df_demo_LET_Available$proxy_SES=="Eligible",1,0),use = "complete.obs")
```


# Are the visual measures different across the FRPM eligible and inelligible groups? 

 - in Kinder they are not the LEtter task is in g1 and MP also shows up as significant by grade 1
```{r}
# MEP-LET

df_frmp_LET<- df_demo_LET_Available %>% filter(!(grade == 2| grade == 3)) %>% filter(!(proxy_SES == "NotReported"))  
df_frmp_LET %>% group_by(proxy_SES) %>% get_summary_stats(LetAbilitySS, type = "mean_sd")
df_frmp_LET$proxy_SES <- factor(df_frmp_LET$proxy_SES) #, levels = rev(levels(df_frmp_LET$proxy_SES)))

summary_stats <- df_frmp_LET %>%
  group_by( proxy_SES) %>%
  get_summary_stats(LetAbilitySS, type = "mean_sd")

# Count sample size by group
sample_sizes <- df_frmp_LET %>% group_by(grade) %>% 
  count(proxy_SES) %>%
  rename(SampleSize = n)

# Merge summary statistics and sample sizes
summary_stats <- merge(summary_stats, sample_sizes, by = "proxy_SES")

bxp <- ggboxplot(
  df_frmp_LET, x = "proxy_SES", y = "LetAbilitySS", 
  ylab = "MEP-L (ø)", xlab = "",fill ="proxy_SES" ) + 
  facet_grid(~grade,labeller = labeller(grade = c(
    "0" = "K",
    "1" = "1"))) + geom_text(
     data = sample_sizes %>% mutate(ordered_group = factor(proxy_SES, levels = c("Eligible", "Ineligible"))),
    aes(label = paste("n =", SampleSize),y=rep(-4.3,4), x = ordered_group),
    size = 5
  ) + scale_fill_manual(values = c("skyblue", "slategray"))  + scale_x_discrete(labels = c( "Eligible", "Ineligible")) + theme(axis.text.x = element_text(size = 18),  # Adjust font size of x-axis labels
        axis.text.y = element_text(size = 18), axis.title.y.left = element_text(size = 18), legend.position = "none", strip.text = element_text(size = 18),  #  Increase facet label (panel heading) font size
    text = element_text(size = 18)) # Increase overall text size

bxp

#Identify outliers 
df_frmp_LET %>%
  group_by(proxy_SES) %>%
  identify_outliers(LetAbilitySS)

# Compute Shapiro wilk test by goups

df_frmp_LET %>%
  group_by(proxy_SES) %>%
  shapiro_test(LetAbilitySS)
# Draw a qq plot by group
ggqqplot(df_frmp_LET, x = "LetAbilitySS", facet.by = "proxy_SES")
#Testing for equal variance assumption to run the test
df_frmp_LET %>% levene_test(LetAbilitySS ~ proxy_SES) 
df_frmp_LET %>%  cohens_d(LetAbilitySS ~ proxy_SES, var.equal = TRUE)
#Levene's test is not significant - therefore we can proceed with the independent samplet t-test otherwise resort to the Weltch ttest


stat.test2 <- df_frmp_LET %>% group_by(grade) %>%
  t_test(LetAbilitySS ~ proxy_SES, var.equal = FALSE) %>%
  add_significance()
stat.test2

# Since there are 9 comparisons in total = we adjsut for those multiple t-tests
alpha = 0.05/12
# Add adjusted p-values to the figure
stat.test_adjusted <- stat.test2 %>%
  mutate(p.adj = p/12) %>% 
  mutate(asteriks = ifelse(p <= alpha, "*", "ns")) %>%
  mutate(cohen.d = lapply(split(df_frmp_LET, df_frmp_LET$grade), function(x) {
    cohen.d(LetAbilitySS ~ proxy_SES, data = x)$cohen.d
  })) %>%
  mutate(cohen.d_effect = unlist(lapply(cohen.d, function(x) round(x[2], 3)))) %>%
  add_xy_position(x = "proxy_SES", dodge = 2.5)

stat.test <- stat.test_adjusted %>% add_xy_position(x = "proxy_SES",dodge =2.5)
bxp + 
  stat_pvalue_manual(stat.test,label="asteriks"  ,tip.length = 0.003,position = position_dodge(0.8), color = "gray2") +
  stat_pvalue_manual(
    stat.test_adjusted, label = "cohen.d_effect", tip.length = 0.03,
    position = position_dodge(1.2), color = "gray2", vjust = 1.5
  )

pValues_tracker =stat.test2$p; 

```
```{r}
# MEP-PSE
df_frpm_PSE<- df_demo_PSE_Available %>% filter(grade !=3 & grade !=2) %>% filter(!(proxy_SES == "NotReported"))  
df_frpm_PSE %>% group_by(proxy_SES) %>% get_summary_stats(pseAbilitySS, type = "mean_sd")
df_frpm_PSE$proxy_SES <- factor(df_frpm_PSE$proxy_SES)
# Count sample size by group
sample_sizes <- df_frpm_PSE %>% group_by(grade) %>% 
  count(proxy_SES) %>%
  rename(SampleSize = n)

bxp <- ggboxplot(
  df_frpm_PSE, x = "proxy_SES", y = "pseAbilitySS", 
  ylab = "MEP-P (ø)", xlab = "", fill = "proxy_SES"
  ) + facet_grid(~grade,labeller = labeller(grade = c(
    "0" = "K",
    "1" = "1"))) + geom_text(
     data = sample_sizes %>% mutate(ordered_group = factor(proxy_SES, levels = c("Eligible", "Ineligible"))),
    aes(label = paste("n =", SampleSize),y=rep(-4.3,4), x = ordered_group),
    size = 5
  ) + scale_fill_manual(values = c("skyblue", "slategray")) + theme(axis.text.x = element_text(size = 18),  # Adjust font size of x-axis labels
        axis.text.y = element_text(size = 18), axis.title.y.left = element_text(size = 18), legend.position = "none", strip.text = element_text(size = 18),  #  Increase facet label (panel heading) font size
    text = element_text(size = 18)) # Increase overall text size

bxp

#Identify outliers 
df_frpm_PSE %>%
  group_by(proxy_SES) %>%
  identify_outliers(pseAbilitySS)

# Compute Shapiro wilk test by goups

df_frpm_PSE %>%
  group_by(proxy_SES) %>%
  shapiro_test(pseAbilitySS)
# Draw a qq plot by group
ggqqplot(df_frpm_PSE, x = "pseAbilitySS", facet.by = "proxy_SES")
#Testing for equal variance assumption to run the test
df_frpm_PSE %>% levene_test(pseAbilitySS ~ proxy_SES) 
df_frpm_PSE %>%  cohens_d(pseAbilitySS ~ proxy_SES, var.equal = TRUE)

#Levene's test is not significant - therefore we can proceed with the independent samplet t-test otherwise resort to the Weltch ttest
stat.test2 <- df_frpm_PSE %>% group_by(grade) %>%
  t_test(pseAbilitySS ~ proxy_SES, var.equal = FALSE) %>%
  add_significance()
stat.test2


# Since there are 9 comparisons in total = we adjsut for those multiple t-tests
alpha = 0.05/12
# Add adjusted p-values to the figure
stat.test_adjusted <- stat.test2 %>%
  mutate(p.adj = p/12) %>% 
  mutate(asteriks = ifelse(p <= alpha, "*", "ns")) %>%
  mutate(cohen.d = lapply(split(df_frpm_PSE, df_frpm_PSE$grade), function(x) {
    cohen.d(pseAbilitySS ~ proxy_SES, data = x)$cohen.d
  })) %>%
  mutate(cohen.d_effect = unlist(lapply(cohen.d, function(x) round(x[2], 3)))) %>%
  add_xy_position(x = "proxy_SES", dodge = 2.5)
  

stat.test <- stat.test_adjusted %>% add_xy_position(x = "proxy_SES",dodge =2.5)
bxp + 
  stat_pvalue_manual(stat.test,label="asteriks"  ,tip.length = 0.003,position = position_dodge(0.8), color = "gray2") +
  stat_pvalue_manual(
    stat.test_adjusted, label = "cohen.d_effect", tip.length = 0.03,
    position = position_dodge(1.2), color = "gray2", vjust = 1.5
  )

pValues_tracker = c(pValues_tracker, stat.test2$p);

```
```{r}
# MP 
df_frpm_GMC<- df_demo_GMC_Available %>% filter(grade !=3 & grade !=2) %>% filter(!(proxy_SES == "NotReported"))  
df_frpm_GMC %>% group_by(proxy_SES) %>% get_summary_stats(MPabilitySS, type = "mean_sd")
df_frpm_GMC$proxy_SES <- factor(df_frpm_GMC$proxy_SES)
# Count sample size by group
sample_sizes <- df_frpm_GMC %>% group_by(grade) %>% 
  count(proxy_SES) %>%
  rename(SampleSize = n)

bxp <- ggboxplot(
  df_frpm_GMC, x = "proxy_SES", y = "MPabilitySS", 
  ylab = "Motion (ø)", xlab = " ", fill = "proxy_SES"
  ) + facet_grid(~grade,labeller = labeller(grade = c(
    "0" = "K",
    "1" = "1"))) + geom_text(
     data = sample_sizes %>% mutate(ordered_group = factor(proxy_SES, levels = c("Eligible", "Ineligible"))),
    aes(label = paste("n =", SampleSize),y=rep(-4.3,4), x = ordered_group),
    size = 5
  ) + scale_fill_manual(values = c("skyblue", "slategray")) + theme(axis.text.x = element_text(size = 18),  # Adjust font size of x-axis labels
        axis.text.y = element_text(size = 18), axis.title.y.left = element_text(size = 18), legend.position = "none", strip.text = element_text(size = 18),  #  Increase facet label (panel heading) font size
    text = element_text(size = 18)) # Increase overall text size
bxp

#Identify outliers 
df_frpm_GMC %>%
  group_by(proxy_SES) %>%
  identify_outliers(MPabilitySS)

# Compute Shapiro wilk test by goups

df_frpm_GMC %>%
  group_by(proxy_SES) %>%
  shapiro_test(MPabilitySS)
# Draw a qq plot by group
ggqqplot(df_frpm_GMC, x = "MPabilitySS", facet.by = "proxy_SES")
#Testing for equal variance assumption to run the test
df_frpm_GMC %>% levene_test(MPabilitySS ~ proxy_SES) 

df_frpm_GMC %>%  cohens_d(MPabilitySS ~ proxy_SES, var.equal = TRUE)

#Levene's test is not significant - therefore we can proceed with the independent samplet t-test otherwise resort to the Weltch ttest
stat.test2 <- df_frpm_GMC %>% group_by(grade) %>%
  t_test(MPabilitySS ~ proxy_SES, var.equal = FALSE) %>%
  add_significance()
stat.test2

# Since there are 9 comparisons in total = we adjsut for those multiple t-tests
alpha = 0.05/6
# Add adjusted p-values to the figure
stat.test_adjusted <- stat.test2 %>%
  mutate(p.adj = p/12) %>% 
  mutate(asteriks = ifelse(p <= alpha, "*", "ns")) %>%
   mutate(cohen.d = lapply(split(df_frpm_GMC, df_frpm_GMC$grade), function(x) {
    cohen.d(MPabilitySS ~ proxy_SES, data = x)$cohen.d
  })) %>%
  mutate(cohen.d_effect = unlist(lapply(cohen.d, function(x) round(x[2], 3)))) %>%
  add_xy_position(x = "proxy_SES", dodge = 2.5)

stat.test <- stat.test_adjusted %>% add_xy_position(x = "proxy_SES",dodge =2.5)
bxp + 
  stat_pvalue_manual(stat.test,label="asteriks"  ,tip.length = 0.003,position = position_dodge(0.8), color = "gray2") +
  stat_pvalue_manual(
    stat.test_adjusted, label = "cohen.d_effect", tip.length = 0.03,
    position = position_dodge(1.2), color = "gray2", vjust = 1.5
  )

pValues_tracker = c(pValues_tracker, stat.test2$p);

```
```{r}
#False positive rate detection 
 
fdrs<-p.adjust(pValues_tracker, method="BH")
fdrs
```

## Do children with different primary language, perform differently in the visual measures? 

```{r}
# MEP-LET
df_primL_LET<- df_demo_LET_Available %>% filter(!(grade == 2| grade == 3))
df_primL_LET %>% group_by(ELStatus) %>% get_summary_stats(LetAbilitySS, type = "mean_sd")

# Count sample size by group
sample_sizes <- df_primL_LET %>% group_by(grade) %>% 
  count(ELStatus) %>%
  rename(SampleSize = n)

bxp <- df_primL_LET %>% 
  filter(grade!=3) %>%  
   mutate(ELStatus = factor(ELStatus, levels = c("EO", "EL"))) %>%
  ggboxplot( x = "ELStatus", y = "LetAbilitySS", 
  ylab = "MEP-L (ø)", xlab = "" , fill = "ELStatus"
  ) + facet_grid(~grade, labeller = labeller(grade = c(
    "0" = "K",
    "1" = "1"))) + geom_text(
     data = sample_sizes %>% mutate(ordered_group = factor(ELStatus, levels = c("EL", "EO"))),
    aes(label = paste("n =", SampleSize),y=rep(-4.3,4), x = ordered_group),
    size = 5
  ) + scale_fill_manual(values = c("ivory3","darkslategray")) + scale_x_discrete(labels = c( "English", "Spanish")) +theme(axis.text.x = element_text(size = 18),  # Adjust font size of x-axis labels
        axis.text.y = element_text(size = 18), axis.title.y.left = element_text(size = 18), legend.position = "none", strip.text = element_text(size = 18),  #  Increase facet label (panel heading) font size
    text = element_text(size = 18)) # Increase overall text size
bxp

#Identify outliers 
df_primL_LET %>%
  group_by(ELStatus) %>%
  identify_outliers(LetAbilitySS)

# Compute Shapiro wilk test by groups

df_primL_LET %>%
  group_by(ELStatus) %>%
  shapiro_test(LetAbilitySS)
# Draw a qq plot by group
ggqqplot(df_primL_LET, x = "LetAbilitySS", facet.by = "ELStatus")
#Testing for equal variance assumption to run the test
df_primL_LET %>% levene_test(LetAbilitySS ~ ELStatus) 

#Levene's test is not significant - therefore we can proceed with the independent samplet t-test otherwise resort to the Weltch ttest
stat.test2 <- df_primL_LET %>% #filter((langinst =="en" & ELStatus =="EL") | (langinst =="es" & ELStatus =="EL") |  (langinst =="en" & ELStatus =="EO"))  %>% #
  #filter((taskInst =="es" & ELStatus =="EL") | (taskInst =="en" & ELStatus =="EO")) %>% 
  filter(grade!=3) %>%
  group_by(grade) %>%
  t_test(LetAbilitySS ~ ELStatus, var.equal = FALSE) %>%  #summarise(mean_diff = mean(LetAbilitySS[ELStatus == "EO"]) - mean(LetAbilitySS[ELStatus == "EL"]),
  ##          SE = sqrt(var(LetAbilitySS[ELStatus == "EO"])/length(LetAbilitySS[ELStatus == "EO"]) + var(LetAbilitySS[ELStatus == "EL"])/length(LetAbilitySS[ELStatus == "EL"])))
  add_significance()
stat.test2

df_primL_LET %>%  cohens_d(LetAbilitySS ~ ELStatus, var.equal = TRUE)

# Since there are 9 comparisons in total = we adjust for those multiple t-tests
alpha = 0.05/12
# Add adjusted p-values to the figure
stat.test_adjusted <- stat.test2 %>%
  mutate(p.adj = p/12) %>% 
  mutate(asteriks = ifelse(p <= alpha, "*", "ns")) %>%
 mutate(cohen.d = lapply(split(df_primL_LET, df_primL_LET$grade), function(x) {
    cohen.d(LetAbilitySS ~ ELStatus, data = x)$cohen.d
  })) %>%
  mutate(cohen.d_effect = unlist(lapply(cohen.d, function(x) round(x[2], 3)))) %>%
  add_xy_position(x = "ELStatus", dodge = 2.5)


stat.test <- stat.test_adjusted %>% add_xy_position(x = "ELStatus",dodge =2.5)
bxp + 
  stat_pvalue_manual(stat.test,label="asteriks"  ,tip.length = 0.003,position = position_dodge(0.8), color = "gray2") +
  stat_pvalue_manual(
    stat.test_adjusted, 
    label = "cohen.d_effect", 
    color = "gray2", 
    vjust = 1.5,
    position = position_dodge(0.8)
  )

pValues_tracker2 = stat.test2$p;

```
```{r}
# MEP-PSE
df_primL_PSE<- df_demo_PSE_Available %>% filter(!(grade == 2| grade == 3))
df_primL_PSE %>% group_by(ELStatus) %>% get_summary_stats(pseAbilitySS, type = "mean_sd")
# Count sample size by group
sample_sizes <- df_primL_PSE %>% group_by(grade) %>% 
  count(ELStatus) %>%
  rename(SampleSize = n)
bxp <- df_primL_PSE %>% 
   mutate(ELStatus = factor(ELStatus, levels = c("EO", "EL"))) %>%
  filter(grade!=3) %>%
  ggboxplot(x = "ELStatus", y = "pseAbilitySS", 
  ylab = "MEP-P (ø)", xlab = " ", fill = "ELStatus"
  ) + facet_grid(~grade, labeller = labeller(grade = c(
    "0" = "K",
    "1" = "1"))) + geom_text(
     data = sample_sizes %>% mutate(ordered_group = factor(ELStatus, levels = c("EL", "EO"))),
    aes(label = paste("n =", SampleSize),y=rep(-4.3,4), x = ordered_group),
    size = 5
  ) + scale_fill_manual(values = c("ivory3", "darkslategray")) + scale_x_discrete(labels = c( "English", "Spanish")) +theme(axis.text.x = element_text(size = 18),  # Adjust font size of x-axis labels
        axis.text.y = element_text(size = 18), axis.title.y.left = element_text(size = 18), legend.position = "none", strip.text = element_text(size = 18),  #  Increase facet label (panel heading) font size
    text = element_text(size = 18)) # Increase overall text size
bxp

#Identify outliers 
df_primL_PSE %>%
  group_by(ELStatus) %>%
  identify_outliers(pseAbilitySS)

# Compute Shapiro wilk test by goups

df_primL_PSE %>%
  group_by(ELStatus) %>%
  shapiro_test(pseAbilitySS)
# Draw a qq plot by group
ggqqplot(df_primL_PSE, x = "pseAbilitySS", facet.by = "ELStatus")
#Testing for equal variance assumption to run the test
df_primL_PSE %>% levene_test(pseAbilitySS ~ ELStatus) 

#Levene's test is not significant - therefore we can proceed with the independent samplet t-test otherwise resort to the Weltch ttest
stat.test2 <- df_primL_PSE %>% #filter((langinst =="en" & ELStatus =="EL") | (langinst =="es" & ELStatus =="EL") |  (langinst =="en" & ELStatus =="EO"))  %>%  
  #filter((taskInst =="es" & ELStatus =="EL") | (taskInst =="en" & ELStatus =="EO"))  %>% 
  filter(grade!=3) %>%
  group_by(grade) %>%
  t_test(pseAbilitySS ~ ELStatus, var.equal = FALSE) %>%
  add_significance()
stat.test2

df_demo_PSE_Available %>%  cohens_d(pseAbilitySS ~ ELStatus, var.equal = TRUE)


# Since there are 9 comparisons in total = we adjsut for those multiple t-tests
alpha = 0.05/12
# Add adjusted p-values to the figure

stat.test_adjusted <- stat.test2 %>%
  mutate(p.adj = p/12) %>% 
  mutate(asteriks = ifelse(p <= alpha, "*", "ns")) %>%
   mutate(cohen.d = lapply(split(df_primL_PSE, df_primL_PSE$grade), function(x) {
    cohen.d(pseAbilitySS ~ ELStatus, data = x)$cohen.d
  })) %>%
  mutate(cohen.d_effect = unlist(lapply(cohen.d, function(x) round(x[2], 3)))) %>%
  add_xy_position(x = "ELStatus", dodge = 2.5)

stat.test_simplified <- stat.test_adjusted %>%
  select(grade, group1, group2, cohen.d_effect, y.position)

stat.test <- stat.test_adjusted %>% add_xy_position(x = "ELStatus",dodge =2.5)
bxp + 
  stat_pvalue_manual(stat.test,label="asteriks"  ,tip.length = 0.003,position = position_dodge(0.8), color = "gray2") +
  stat_pvalue_manual(
    stat.test_simplified, 
    label = "cohen.d_effect", 
    color = "gray2", 
    vjust = 1.5,
    position = position_dodge(0.8)
  )

#pValues_tracker2 = c(pValues_tracker2, stat.test2$p);

```
```{r}
# MP 
df_primL_GMC<- df_demo_GMC_Available %>% filter(!(grade == 2| grade == 3))
df_primL_GMC %>% group_by(ELStatus) %>% get_summary_stats(MPabilitySS, type = "mean_sd")

sample_sizes <- df_primL_GMC %>% group_by(grade) %>% 
  count(ELStatus) %>%
  rename(SampleSize = n) %>% mutate(ELStatus = factor(ELStatus, levels = c("EO", "EL")))
sample_sizes <- sample_sizes %>%
  mutate(new_order = case_when(
    grade == 0 & ELStatus == "EO" ~ 1,
    grade == 0 & ELStatus == "EL" ~ 2,
    grade == 1 & ELStatus == "EO" ~ 3,
    grade == 1 & ELStatus == "EL" ~ 4
  )) %>%
  arrange(new_order) %>%
  select(-new_order)

bxp <- ggboxplot(
  df_primL_GMC, x = "ELStatus", y = "MPabilitySS", 
  ylab = "Motion (ø)", xlab = "", fill = "ELStatus"
  ) + facet_grid(~grade, labeller = labeller(grade = c(
    "0" = "K",
    "1" = "1"))) + geom_text(
     data = sample_sizes, aes(label = paste("n =", SampleSize),y=rep(-4.3,4), x= rep(c(1,2),2)),
    size = 5)  + scale_fill_manual(values = c( "ivory3", "darkslategray")) + scale_x_discrete(labels = c( "English", "Spanish")) +theme(axis.text.x = element_text(size = 18),  # Adjust font size of x-axis labels
        axis.text.y = element_text(size = 18), axis.title.y.left = element_text(size = 18), legend.position = "none", strip.text = element_text(size = 18),  #  Increase facet label (panel heading) font size
    text = element_text(size = 18)) # Increase overall text size
bxp


#Levene's test is not significant - therefore we can proceed with the independent samplet t-test otherwise resort to the Weltch ttest
stat.test2 <- df_primL_GMC %>% group_by(grade) %>%
  t_test(MPabilitySS ~ ELStatus, var.equal = TRUE) %>%
  add_significance()
stat.test2

df_primL_GMC %>%  cohens_d(MPabilitySS ~ ELStatus, var.equal = TRUE)

# Since there are 9 comparisons in total = we adjsut for those multiple t-tests
alpha = 0.05/12
# Add adjusted p-values to the figure
stat.test_adjusted <- stat.test2 %>%
  mutate(p.adj = p/12) %>% 
  mutate(asteriks = ifelse(p <= alpha, "*", "ns")) %>%
   mutate(cohen.d = lapply(split(df_primL_GMC, df_primL_GMC$grade), function(x) {
    cohen.d(MPabilitySS ~ ELStatus, data = x)$cohen.d
  })) %>%
  mutate(cohen.d_effect = unlist(lapply(cohen.d, function(x) round(x[2], 3)))) %>%
  add_xy_position(x = "ELStatus", dodge = 2.5)

stat.test_simplified <- stat.test_adjusted %>%
  select(grade, group1, group2, cohen.d_effect, y.position)

stat.test <- stat.test_adjusted %>% add_xy_position(x = "ELStatus",dodge =2.5)
bxp + 
  stat_pvalue_manual(stat.test,label="asteriks"  ,tip.length = 0.003,position = position_dodge(0.8), color = "gray2") +
  stat_pvalue_manual(
    stat.test_simplified, 
    label = "cohen.d_effect", 
    color = "gray2", 
    vjust = 1.5,
    position = position_dodge(0.8)
  )



```

### Supplementary Figure 2: Check how the reading measures are across SES and ELStatus 

```{r}
# MEP-LET
df_frpm_LET<- df_demo_LET_Available %>% filter(!(grade == 2| grade == 3)) %>% filter(!(proxy_SES == "NotReported")) %>%  mutate(across(c(wcj_lwi_ss,wcj_wa_ss,wcj_spelling_ss,wcm_lwi_ss), as.numeric)) %>%
  mutate(reading_outcome = wcj_lwi_ss) #if_else(grade == 0,if_else(!is.na(wcm_lwi_ss), wcm_lwi_ss, wcj_lwi_ss), wcj_lwi_ss))
df_frpm_LET$proxy_SES <- factor(df_frpm_LET$proxy_SES)

# Count sample size by group
sample_sizes <- df_frpm_LET %>% group_by(grade) %>% filter(complete.cases(wcj_lwi_ss)) %>%
  count(proxy_SES) %>%
  rename(SampleSize = n)

bxp <- ggboxplot(
  df_frpm_LET, x = "proxy_SES", y = "reading_outcome", 
  ylab = "Reading outcome (LWI)", xlab = "", fill = "proxy_SES"
  ) + facet_grid(~grade,  labeller = labeller(grade = c(
    "0" = "K",
    "1" = "1")))  + geom_text(
    data = sample_sizes,
    aes(label = paste("n =", SampleSize),y=rep(-2.9,4), x= rep(c(1,2),2)),
    size = 5
  ) + scale_fill_manual(values = c("skyblue", "slategray")) + theme(axis.text.x = element_text(size = 18),  # Adjust font size of x-axis labels
        axis.text.y = element_text(size = 18), axis.title.y.left = element_text(size = 18), legend.position = "none", strip.text = element_text(size = 18),  #  Increase facet label (panel heading) font size
    text = element_text(size = 18)) # Increase overall text size

bxp

stat.test2 <- df_frpm_LET %>% group_by(grade) %>%
  t_test(reading_outcome ~ proxy_SES, var.equal = FALSE) %>%
  add_significance()
stat.test2

# Since there are 9 comparisons in total = we adjust for those multiple t-tests
alpha = 0.05/12
# Add adjusted p-values to the figure
stat.test_adjusted <- stat.test2 %>%
  mutate(p.adj = p/12) %>% mutate(asteriks = ifelse(stat.test2$p <= alpha, "*", ifelse(stat.test2$p <= 0.001,"***", "ns"))) %>%
   mutate(cohen.d = lapply(split(df_frpm_LET, df_frpm_LET$grade), function(x) {
    cohen.d(reading_outcome ~ proxy_SES, data = x)$cohen.d
  })) %>%
  mutate(cohen.d_effect = unlist(lapply(cohen.d, function(x) round(x[2], 3)))) %>%
  add_xy_position(x = "proxy_SES", dodge = 2.5)

stat.test <- stat.test_adjusted %>% add_xy_position(x = "proxy_SES",dodge =2.5)
bxp + 
  stat_pvalue_manual(stat.test,label="asteriks"  ,tip.length = 0.003,position = position_dodge(0.8), color = "gray2") +
  stat_pvalue_manual(
    stat.test_adjusted, label = "cohen.d_effect", tip.length = 0.03,
    position = position_dodge(1.2), color = "gray2", vjust = 1.5
  )

```
```{r}
# MEP-PSE
df_frpm_PSE<- df_demo_PSE_Available %>% filter(!(grade == 2| grade == 3)) %>% filter(!(proxy_SES == "NotReported")) %>% mutate(across(c(wcj_lwi_ss,wcj_wa_ss,wcj_spelling_ss,wcm_lwi_ss), as.numeric)) %>%
  mutate(reading_outcome =  wcj_lwi_ss) #if_else(grade == 0,if_else(!is.na(wcm_lwi_ss), wcm_lwi_ss, wcj_lwi_ss), wcj_lwi_ss))
df_frpm_PSE$proxy_SES <- factor(df_frpm_PSE$proxy_SES)

# Count sample size by group
sample_sizes <- df_frpm_PSE %>% group_by(grade) %>% filter(complete.cases(wcj_lwi_ss)) %>%
  count(proxy_SES) %>%
  rename(SampleSize = n)

bxp <- ggboxplot(
  df_frpm_PSE, x = "proxy_SES", y = "reading_outcome", 
  ylab = "Reading outcome (LWI)", xlab = "", fill = "proxy_SES"
  ) + facet_grid(~grade,  labeller = labeller(grade = c(
    "0" = "K",
    "1" = "1")))  + geom_text(
    data = sample_sizes,
    aes(label = paste("n =", SampleSize),y=rep(-2.9,4), x= rep(c(1,2),2)),
    size = 5
  ) +  scale_fill_manual(values = c("skyblue", "slategray")) + theme(axis.text.x = element_text(size = 18),  # Adjust font size of x-axis labels
        axis.text.y = element_text(size = 18), axis.title.y.left = element_text(size = 18), legend.position = "none", strip.text = element_text(size = 18),  #  Increase facet label (panel heading) font size
    text = element_text(size = 18)) # Increase overall text size

bxp

stat.test2 <- df_frpm_PSE %>% group_by(grade) %>%
  t_test(reading_outcome ~ proxy_SES, var.equal = FALSE) %>%
  add_significance()
stat.test2

# Since there are 9 comparisons in total = we adjsut for those multiple t-tests
alpha = 0.05/12
# Add adjusted p-values to the figure
stat.test_adjusted <- stat.test2 %>%
  mutate(p.adj = p/12) %>% mutate(asteriks = ifelse(stat.test2$p <= alpha, "*", ifelse(stat.test2$p <= 0.001,"***", "ns"))) %>%
   mutate(cohen.d = lapply(split(df_frpm_PSE, df_frpm_PSE$grade), function(x) {
    cohen.d(reading_outcome ~ proxy_SES, data = x)$cohen.d
  })) %>%
  mutate(cohen.d_effect = unlist(lapply(cohen.d, function(x) round(x[2], 3)))) %>%
  add_xy_position(x = "proxy_SES", dodge = 2.5)

stat.test <- stat.test_adjusted %>% add_xy_position(x = "proxy_SES",dodge =2.5)
bxp + 
  stat_pvalue_manual(stat.test,label="asteriks"  ,tip.length = 0.003,position = position_dodge(0.8), color = "gray2") +
  stat_pvalue_manual(
    stat.test_adjusted, label = "cohen.d_effect", tip.length = 0.03,
    position = position_dodge(1.2), color = "gray2", vjust = 1.5
  )
```
```{r}
# Motion
df_frpm_GMC<- df_demo_GMC_Available %>% filter(!(grade == 2| grade == 3)) %>% filter(!(proxy_SES == "NotReported")) %>% mutate(across(c(wcj_lwi_ss,wcj_wa_ss,wcj_spelling_ss,wcm_lwi_ss), as.numeric)) %>%
  mutate(reading_outcome = wcj_lwi_ss) #if_else(grade == 0,if_else(!is.na(wcm_lwi_ss), wcm_lwi_ss, wcj_lwi_ss), wcj_lwi_ss))
df_frpm_GMC$proxy_SES <- factor(df_frpm_GMC$proxy_SES)
# Count sample size by group
sample_sizes <- df_frpm_GMC %>% group_by(grade) %>% filter(complete.cases(wcj_lwi_ss)) %>%
  count(proxy_SES) %>%
  rename(SampleSize = n)

bxp <- ggboxplot(
  df_frpm_GMC, x = "proxy_SES", y = "reading_outcome", 
  ylab = "Reading outcome (LWI)", xlab = "", fill = "proxy_SES") + facet_grid(~grade, labeller = labeller(grade = c(
    "0" = "K",
    "1" = "1")))  + geom_text(
    data = sample_sizes,
    aes(label = paste("n =", SampleSize),y=rep(-2.9,4), x= rep(c(1,2),2)),
    size = 5
  ) +  scale_fill_manual(values = c("skyblue", "slategray")) + theme(axis.text.x = element_text(size = 18),  # Adjust font size of x-axis labels
        axis.text.y = element_text(size = 18), axis.title.y.left = element_text(size = 18), legend.position = "none", strip.text = element_text(size = 18),  #  Increase facet label (panel heading) font size
    text = element_text(size = 18)) # Increase overall text size

bxp

stat.test2 <- df_frpm_GMC %>% group_by(grade) %>%
  t_test(reading_outcome ~ proxy_SES, var.equal = FALSE) %>%
  add_significance()
stat.test2

# Since there are 9 comparisons in total = we adjsut for those multiple t-tests
alpha = 0.05/12
# Add adjusted p-values to the figure
stat.test_adjusted <- stat.test2 %>%
  mutate(p.adj = p/12) %>% mutate(asteriks = ifelse(stat.test2$p <= alpha, "*", ifelse(stat.test2$p <= 0.001,"***", "ns"))) %>%
   mutate(cohen.d = lapply(split(df_frpm_GMC, df_frpm_GMC$grade), function(x) {
    cohen.d(reading_outcome ~ proxy_SES, data = x)$cohen.d
  })) %>%
  mutate(cohen.d_effect = unlist(lapply(cohen.d, function(x) round(x[2], 3)))) %>%
  add_xy_position(x = "proxy_SES", dodge = 2.5)

stat.test <- stat.test_adjusted %>% add_xy_position(x = "proxy_SES",dodge =2.5)
bxp + 
  stat_pvalue_manual(stat.test,label="asteriks"  ,tip.length = 0.003,position = position_dodge(0.8), color = "gray2") +
  stat_pvalue_manual(
    stat.test_adjusted, label = "cohen.d_effect", tip.length = 0.03,
    position = position_dodge(1.2), color = "gray2", vjust = 1.5
  )

```
```{r}
# MEP-LET
df_primL_LET <- df_primL_LET %>% mutate(across(c(wcj_lwi_ss,wcj_wa_ss,wcj_spelling_ss,wcm_lwi_ss), as.numeric)) %>%
  mutate(reading_outcome = wcj_lwi_ss)
# Count sample size by group
sample_sizes <- df_primL_LET %>% group_by(grade) %>% 
  count(ELStatus) %>%
  rename(SampleSize = n)
bxp <- df_primL_LET %>% 
   mutate(ELStatus = factor(ELStatus, levels = c("EO", "EL"))) %>%
  filter(grade!=3) %>%
  ggboxplot(x = "ELStatus", y = "reading_outcome", 
  ylab = "Reading_outcome (LWI)", xlab = " ", fill = "ELStatus"
  ) + facet_grid(~grade, labeller = labeller(grade = c(
    "0" = "K",
    "1" = "1"))) + geom_text(
    data = sample_sizes,
    aes(label = paste("n =", SampleSize),y=rep(-4.3,4), x= rep(c(1,2),2)),
    size =5
  ) + scale_fill_manual(values = c("ivory3", "darkslategray")) + scale_x_discrete(labels = c( "English", "Spanish"))  + theme(axis.text.x = element_text(size = 18),  # Adjust font size of x-axis labels
        axis.text.y = element_text(size = 18), axis.title.y.left = element_text(size = 18), legend.position = "none", strip.text = element_text(size = 18),  #  Increase facet label (panel heading) font size
    text = element_text(size = 18)) # Increase overall text size
bxp


#Levene's test is not significant - therefore we can proceed with the independent samplet t-test otherwise resort to the Weltch ttest
stat.test2 <- df_primL_LET %>% #filter((langinst =="en" & ELStatus =="EL") | (langinst =="es" & ELStatus =="EL") |  (langinst =="en" & ELStatus =="EO"))  %>%  
  #filter((taskInst =="es" & ELStatus =="EL") | (taskInst =="en" & ELStatus =="EO"))  %>% 
  filter(grade!=3 | grade!=2) %>%
  group_by(grade) %>%
  t_test(reading_outcome ~ ELStatus, var.equal = FALSE) %>%
  add_significance()
stat.test2

# Since there are 9 comparisons in total = we adjsut for those multiple t-tests
alpha = 0.05/12
# Add adjusted p-values to the figure
stat.test_adjusted <- stat.test2 %>%
  mutate(p.adj = p/12) %>% mutate(asteriks = ifelse(stat.test2$p <= alpha, "*", ifelse(stat.test2$p <= 0.001,"***", "ns"))) %>%
   mutate(cohen.d = lapply(split(df_primL_LET, df_primL_LET$grade), function(x) {
    cohen.d(reading_outcome ~ ELStatus, data = x)$cohen.d
  })) %>%
  mutate(cohen.d_effect = unlist(lapply(cohen.d, function(x) round(x[2], 3)))) %>%
  add_xy_position(x = "ELStatus", dodge = 2.5)

stat.test <- stat.test_adjusted %>% add_xy_position(x = "ELStatus",dodge =2.5)
bxp + 
  stat_pvalue_manual(stat.test,label="asteriks"  ,tip.length = 0.003,position = position_dodge(0.8), color = "gray2") +
  stat_pvalue_manual(
    stat.test_adjusted, label = "cohen.d_effect", tip.length = 0.03,
    position = position_dodge(1.2), color = "gray2", vjust = 1.5
  )

pValues_tracker2 = c(pValues_tracker2, stat.test2$p);


```

```{r}
# MEP-PSE
df_primL_PSE <- df_primL_PSE %>% mutate(across(c(wcj_lwi_ss,wcj_wa_ss,wcj_spelling_ss,wcm_lwi_ss), as.numeric)) %>%
  mutate(reading_outcome = wcj_lwi_ss)
# Count sample size by group
sample_sizes <- df_primL_PSE %>% group_by(grade) %>% 
  count(ELStatus) %>%
  rename(SampleSize = n) %>% mutate(ELStatus = factor(ELStatus, levels = c("EO", "EL")))
sample_sizes <- sample_sizes %>%
  mutate(new_order = case_when(
    grade == 0 & ELStatus == "EO" ~ 1,
    grade == 0 & ELStatus == "EL" ~ 2,
    grade == 1 & ELStatus == "EO" ~ 3,
    grade == 1 & ELStatus == "EL" ~ 4
  )) %>%
  arrange(new_order) %>%
  select(-new_order)
bxp <- df_primL_PSE %>% 
   mutate(ELStatus = factor(ELStatus, levels = c("EO", "EL"))) %>%
  filter(grade!=3) %>%
  ggboxplot(x = "ELStatus", y = "reading_outcome", 
  ylab = "Reading_outcome (LWI)", xlab = " ", fill = "ELStatus"
  ) + facet_grid(~grade, labeller = labeller(grade = c(
    "0" = "K",
    "1" = "1"))) + geom_text(
    data = sample_sizes,
    aes(label = paste("n =", SampleSize),y=rep(-4.3,4), x= rep(c(1,2),2)),
    size = 5
  ) + scale_fill_manual(values = c("ivory3", "darkslategray")) + scale_x_discrete(labels = c( "English", "Spanish")) +theme(axis.text.x = element_text(size = 18),  # Adjust font size of x-axis labels
        axis.text.y = element_text(size = 18), axis.title.y.left = element_text(size = 18), legend.position = "none", strip.text = element_text(size = 18),  #  Increase facet label (panel heading) font size
    text = element_text(size = 18)) # Increase overall text size
bxp


#Levene's test is not significant - therefore we can proceed with the independent samplet t-test otherwise resort to the Weltch ttest
stat.test2 <- df_primL_PSE %>% mutate(ELStatus = factor(ELStatus, levels = c("EL", "EO"))) %>% #filter((langinst =="en" & ELStatus =="EL") | (langinst =="es" & ELStatus =="EL") |  (langinst =="en" & ELStatus =="EO"))  %>%  
  #filter((taskInst =="es" & ELStatus =="EL") | (taskInst =="en" & ELStatus =="EO"))  %>% 
  filter(grade!=3 | grade!=2) %>%
  group_by(grade) %>%
  t_test(reading_outcome ~ ELStatus, var.equal = FALSE) %>%
  add_significance()
stat.test2

# Since there are 9 comparisons in total = we adjsut for those multiple t-tests
alpha = 0.05/12
# Add adjusted p-values to the figure
stat.test_adjusted <- stat.test2 %>%
  mutate(p.adj = p/6) %>% mutate(asteriks = ifelse(stat.test2$p <= alpha, "*", ifelse(stat.test2$p <= 0.001,"***", "ns"))) %>%
   mutate(cohen.d = lapply(split(df_primL_PSE, df_primL_PSE$grade), function(x) {
    cohen.d(reading_outcome ~ ELStatus, data = x)$cohen.d
  })) %>%
  mutate(cohen.d_effect = unlist(lapply(cohen.d, function(x) round(x[2], 3)))) %>%
  add_xy_position(x = "ELStatus", dodge = 2.5)

stat.test <- stat.test_adjusted %>% add_xy_position(x = "ELStatus",dodge =2.5)
bxp + 
  stat_pvalue_manual(stat.test,label="asteriks"  ,tip.length = 0.003,position = position_dodge(0.8), color = "gray2") +
  stat_pvalue_manual(
    stat.test_adjusted, label = "cohen.d_effect", tip.length = 0.03,
    position = position_dodge(1.2), color = "gray2", vjust = 1.5
  )

pValues_tracker2 = c(pValues_tracker2, stat.test2$p);

```
```{r}
# MP 
df_primL_GMC <- df_primL_GMC %>% mutate(across(c(wcj_lwi_ss,wcj_wa_ss,wcj_spelling_ss,wcm_lwi_ss), as.numeric)) %>%
  mutate(reading_outcome = wcj_lwi_ss)
#df_demo_GMC_Available<- df_demo_GMC_Available %>% filter(grade !=3) 
df_primL_GMC %>% group_by(ELStatus) %>% get_summary_stats(MPabilitySS, type = "mean_sd")

sample_sizes <- df_primL_GMC %>% group_by(grade) %>% 
  count(ELStatus) %>%
  rename(SampleSize = n)

bxp <- ggboxplot(
  df_primL_GMC, x = "ELStatus", y = "reading_outcome", 
  ylab = "Reading_outcome (LWI)", xlab = "", fill = "ELStatus"
  ) + facet_grid(~grade, labeller = labeller(grade = c(
    "0" = "K",
    "1" = "1"))) + geom_text(
    data = sample_sizes,
    aes(label = paste("n =", SampleSize),y=rep(-4.3,4), x= rep(c(1,2),2)),
    size = 5
  ) + scale_fill_manual(values = c("darkslategray", "ivory3")) + scale_x_discrete(labels = c(  "Spanish", "English")) +theme(axis.text.x = element_text(size = 18),  # Adjust font size of x-axis labels
        axis.text.y = element_text(size = 18), axis.title.y.left = element_text(size = 18), legend.position = "none", strip.text = element_text(size = 18),  #  Increase facet label (panel heading) font size
    text = element_text(size = 18)) # Increase overall text size
bxp

#Levene's test is not significant - therefore we can proceed with the independent samplet t-test otherwise resort to the Weltch ttest
stat.test2 <- df_primL_GMC %>% group_by(grade) %>%
  t_test(reading_outcome ~ ELStatus, var.equal = TRUE) %>%
  add_significance()
stat.test2

# Since there are 9 comparisons in total = we adjsut for those multiple t-tests
alpha = 0.05/12
# Add adjusted p-values to the figure
stat.test_adjusted <- stat.test2 %>%
  mutate(p.adj = p/6) %>% mutate(asteriks = ifelse(stat.test2$p <= alpha, "*", ifelse(stat.test2$p <= 0.001,"***", "ns"))) %>%
   mutate(cohen.d = lapply(split(df_primL_GMC, df_primL_GMC$grade), function(x) {
    cohen.d(reading_outcome ~ ELStatus, data = x)$cohen.d
  })) %>%
  mutate(cohen.d_effect = unlist(lapply(cohen.d, function(x) round(x[2], 3)))) %>%
  add_xy_position(x = "ELStatus", dodge = 2.5)

stat.test <- stat.test_adjusted %>% add_xy_position(x = "ELStatus",dodge =2.5)
bxp + 
  stat_pvalue_manual(stat.test,label="asteriks"  ,tip.length = 0.003,position = position_dodge(0.8), color = "gray2") +
  stat_pvalue_manual(
    stat.test_adjusted, label = "cohen.d_effect", tip.length = 0.03,
    position = position_dodge(1.2), color = "gray2", vjust = 1.5
  )

pValues_tracker2 = c(pValues_tracker2, stat.test2$p);

```
## False detection rate 

```{r eval= FALSE}
fdrs2<-p.adjust(pValues_tracker2, method="BH")
fdrs2
```